Balance responses in MS

### 2020-01-16 R/msBalance.Rmd
updated msBalance.Rmd to call matlab preprocessing
reorganized manuscript

```{r}
source("../paths.m")
```

@@

# Methods and Results

## Data sources
```{r, message=FALSE, warning=FALSE}
library(vroom)
library(dplyr)
library(stringr)
library(lubridate)
`%contains%` = function(s, pattern) grepl(pattern, s)
`%omits%` = function(s, pattern) !grepl(pattern, s)

@@

This study performed a secondary analysis on existing electromyographic recordings during balance responses in people with and without MS.
Individual subjects were as follows:

>>>>

```{r}
# set path variables
source("../paths.m")

# if csv files do not exist, run matlab batch steps
if (!file.exists("../csv/c.csv")|!file.exists("../csv/m.csv")){
  system(paste(matlabCommand, "'cd ../m; loadData; exit'"))
}
```

@@

# Methods

## Data sources

This study performed a secondary analysis on existing electromyographic recordings during balance responses in people with and without MS.
Clinical and demographic characteristics of the study sample are summarized in Table 1.

```{r, message=FALSE, warning=FALSE}
library(vroom)
library(dplyr)
library(stringr)
library(lubridate)
`%contains%` = function(s, pattern) grepl(pattern, s)
`%omits%` = function(s, pattern) !grepl(pattern, s)

@@

## Assessment protocol
Complete details of the assessment protocol are avaialble in an earlier report.[@huisinga2014] 
Briefly, subjects stood on a set of 2 custom hydraulic force platforms that translated forward together causing backward body sway and the activation of tibialis anterior (TA).
subjects stood with arms folded across the chest, eyes open, with their feet at a fixed heel-to-heel distance of 10cm.
Peak platform displacement was 4 cm, at a rate of 15 cm/s, which elicited feet-in-place balance corrections without stepping.
Surface electromyography was recorded from the right TA of all subjects with two 2.5-cm^2^ surface electrodes placed approximately 2 cm apart, with a ground electrode on the lateral condyle.

## Data processing
Each individual EMG record was registered to the onset of platform motion and resampled.

Individual subjects were as follows:

### 2020-01-16 README.Rmd
reorganized manuscript

---
output: github_document
---

@@

# msBalance

## Overview

msBalance is an R project for analysis of balance responses in people with and without multiple sclerosis (MS).
The source data were created by Jessie Huisinga, Ph.D., and collected by Daniel Peterson, Ph.D.
The analysis code was primarily developed by J. Lucas McKay, Ph.D., M.S.C.R.

## Getting and installing msBalance

### Downloading from GitHub for development

1. Use RStudio to create a local copy of the project
    1. In RStudio, start a new Project: `File > New Project > Version Control > Git`
    1. In the “repository URL” paste: https://github.com/jlucasmckay/msBalance
    1. On OS X, you can make this under `tmp` - this is a temporary file that can be deleted later
    1. If there are files that need to be added to the remote, they can be placed in this folder or dragged / dropped into the GitHub interface.
1. Path locations that need to be read by Matlab and R can be specified in `msBalance/paths.m`.
1. Path locations might need to be modified in `msBalance/m/loadData.m`.
1. The main analysis is in `msBalance/R/msBalance.Rmd`
1. Stage changed files, commit, and push changes with the git panel.

### Navigating a mixture of Matlab and R code

1. The source files are in `.mat` format.
1. There is a mixture of matlab preprocessing code (files in `m/`) into `.csv` files that R can then read (files in `R/`).
1. A directory `csv/` is created to hold these files. The contents of this directory are not automatically loaded to GitHub.

### Running the code

1. Alter `paths.m` to point to the correct data directory.
1. Run the Matlab code to load .mat files and populate patient-level data
    * Run `m/loadData.m` to load source files and populate the `csv/` directory.
    * A future version of this could call the Matlab script from R if useful.
1. "knit" `R/msBalance.Rmd` to produce the study document.

>>>>

---
title: "Balance responses in MS"
author: J.L. McKay
date: "`r format(Sys.time(), '%d %B %Y')`"
output: github_document
always_allow_html: yes
---

@@

# Overview
`msBalance` is an R project for analysis of balance responses in people with and without multiple sclerosis (MS).
The source data were created by Jessie Huisinga, Ph.D., and collected by Daniel Peterson, Ph.D.
The analysis code was primarily developed by J. Lucas McKay, Ph.D., M.S.C.R.

# Use

## What to read and edit
1. The main document is `R/msBalance.Rmd`, which can be knit/interpreted into `R/msBalance.html`, which contains the main text and figures.
1. This document explains how to run the code.

## How to run the code
1. Alter `paths.m` in the main project directory so that: 
    1. `dataDirectory` points at the correct data directory on your local machine. This should contain the files `AllConsolidatedData_CS.mat` and `AllConsolidatedData_MS.mat`.
    1. `matlabCommand` points at the appropriate command-line version of Matlab for data preprocessing. On Dr. McKay's machine, this string is `/Applications/MATLAB_R2019a.app/bin/matlab -nojvm -nodesktop -nosplash -r`.
1. "knit" `R/msBalance.Rmd` to produce `R/msBalance.html`. This will call matlab to create intermediary files if they are missing.

## Notes related to the fact that there is a mixture of Matlab and R code
1. The source files are in `.mat` format.
1. There is a mixture of matlab preprocessing code (files in `m/`) into `.csv` files that R can then read (files in `R/`).
1. A directory `csv/` is created to hold these files. The contents of this directory are not automatically loaded to GitHub.

## How to download from GitHub for development
1. In RStudio, start a new Project: `File > New Project > Version Control > Git`.
1. In the “repository URL” paste: https://github.com/jlucasmckay/msBalance.
1. On OS X, you can make this under `tmp` - this is a temporary file that can be deleted later.
1. If there are files that need to be added to the remote, they can be placed in this folder or dragged / dropped into the GitHub interface.
1. Stage changed files, commit, and push changes with the git panel.
